---
title: "Supplementary script to reproduce all analyses : Financial assessment of of continuous cover forestry in a mixed hardwood forest estate "
author: "Gauthier Ligot"
date: "2026-04-20"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
library(sf)
library(mapview)
library(DT)
library(knitr)
# library(xlsx)
library(ggpubr)
# library(readxl)
# library(ggradar) #not available with newer R version
library(RColorBrewer)
library(openxlsx2)
library(kableExtra)


```

# Preamble
This document and script can be used to reproduce the results provided in the manuscript and supplementary materials. 

```{r}
source("script/tirf.R") # a function to compute the internal rate of return
source("script/myGGplotConfigs.R") # settings for personnalized ggplot layouts
load("data/1_CalcData_en.Rdata")  # inventory data
load("data/3_calcRevenues_en.Rdata") # economic data
source("script/commonParameters.R") # economic parameter
fpalette<- c("#E69F00", "#999999", "#009E73", "#D55E00", "#CC79A7","#333399") # chosen colors for the plots
```

```{r}
tmp_purchase<-estate_purchase%>%
  select(block, purchaseYear = year)

keydates<-inventory_year%>%
  filter(year >= 1992)%>%
  group_by(block)%>%
  filter(year == min(year))%>%
  rename(firstInventory = year)%>%
  left_join(tmp_purchase, by=join_by(block))
```


# Results 
## Forest structure and composition
### Initial situation
```{r}
# Table 1
t1<-dendro_block%>%
  as_tibble()%>%
  group_by(block)%>%
  filter(year>= 1992)%>%
  filter(year == min(year))%>%
  select(block, year, nha, gha, vha, vcha, ghapc_oak, ghapc_frax, ghapc_conif)%>%
  mutate_at(c("ghapc_oak","ghapc_frax","ghapc_conif"), function(x) {x *100})%>%
  ungroup()%>%
  mutate_if(is.numeric, ~round(.,digits=1))%>%
  arrange(block)

colnames(t1) <- c("Block", "Year","Density (/ha)","Basal area (m²/ha)", "Standing volume (m³/ha)", 
                  "Standing value (€/ha)", "%Oak", "%Ash", "%Conif.")

datatable(t1,
            caption="Description of the initial density, basal area, standing volume, standing value and composition of each study forest blocks. The composition is described giving the proportion of basal area accounted for by oak, coniferous species or ash. The initial situation corresponds to the year of the first inventory after the purchase of the forest block and after 1992 when the manager decided to apply Pro Silva principles.",
            filter = c("top"))

wb <- wb_workbook()
wb$add_worksheet("t1_dendro")
wb$add_data("t1_dendro", t1, row_names = FALSE, na = "")
```

### Evolution
```{r}
#### Final situation ----
t2<-dendro_block%>%
  as_tibble()%>%
  group_by(block)%>%
  filter(year>= 1992)%>%
  filter(year == max(year))%>%
  select(block, year, nha, gha, vha, vcha, ghapc_oak, ghapc_frax, ghapc_conif)%>%
  mutate_at(c("ghapc_oak","ghapc_frax","ghapc_conif"), function(x) {x *100})%>%
  ungroup()%>%
  mutate_if(is.numeric, ~round(.,digits=1))%>%
  arrange(block)

colnames(t2) <- c("Block", "Year","Density (/ha)","Basal area (m²/ha)", "Standing volume (m³/ha)", 
                  "Standing value (€/ha)", "%Oak", "%Ash", "%Conif.")

datatable(t2,
            caption="Description of the final density, basal area, standing volume, standing value and composition of each study forest blocks. The composition is described giving the proportion of basal area accounted for by oak, coniferous species or ash. The final situation corresponds to the year of the final inventory, in 2019.",
            filter = c("top"))

wb$add_worksheet("tf_dendro")
wb$add_data("tf_dendro", t2, row_names = FALSE, na = "")
```

```{r}
evo7a<-dendro_block%>%
  ggplot(aes(x=year,y=vha))+ geom_line(aes(col=as.factor(block),group=block))+
  geom_point(aes(col=as.factor(block)))+
  xlab("Year")+ylab("Standing volume (m³/ha)")+labs(col="Block")+
  theme(legend.position = "top")+
  theme(axis.text.x=element_text(angle = 45, hjust = 1,vjust = 1))

evo7b<-dendro_block%>%
  ggplot(aes(x=year,y=vcha/vha))+ geom_line(aes(col=as.factor(block),group=block))+
  geom_point(aes(col=as.factor(block)))+
  xlab("Year")+ylab("Standing value / Standing volume (€/m³)")+
  theme(legend.position = "top")+labs(col="Block")+
  theme(axis.text.x=element_text(angle = 45, hjust = 1,vjust = 1))

evo7<-ggarrange(evo7a, evo7b, labels = c("A", "B"), ncol = 2, nrow = 1,
                common.legend = TRUE, legend = "bottom")

ggsave(evo7,file = "fig/fig2.png", width=10, height = 5)
evo7
```


## Timber yield
```{r}
periodsDico<-dendro_block%>%
  as_tibble()%>%
  select(block,year)%>%
  distinct()%>%
  group_by(block)%>%
  arrange(year)%>%
  mutate(year_i = year,
         year_f = lead(year),
         period = row_number())%>%
  select(-year)%>%
  filter(!is.na(year_f))%>%
  group_by(block, period, year_i, year_f)%>%
  nest()%>%
  mutate(year = map2(.x = year_i, .y = year_f, ~ .x:(.y-1)))%>%
  ungroup()%>%
  select(block, period, year, year_i, year_f)%>%
  unnest(cols = c(block, year, year_i, year_f))

harvested<-cashflows%>%
  filter(year >= 1992)%>%
  filter(category == "Timber sales")%>%
  group_by(block,year)%>%
  summarise(volume = sum(volume),
            net_2019 = sum(net_2019))%>%
  left_join(periodsDico, by = join_by(block, year))%>%
  group_by(block)%>%
  mutate(year_f = ifelse(is.na(year_f),min(year_i,na.rm=T),year_f))%>%
  group_by(block,year_f)%>%
  summarise(volume = sum(volume),
            net_2019 = sum(net_2019))

t2a<-dendro_block%>%
  as_tibble()%>%
  select(-geom)%>%
  select(block,block,block_area_ha,year,nha,gha,vha)%>%
  left_join(harvested,join_by(block,year==year_f))%>%
  mutate(vha_harvested = volume/block_area_ha,
         vha_harvested = ifelse(is.na(vha_harvested),0,vha_harvested))%>%
  group_by(block)%>%
  arrange(year)%>%
  mutate(year_i = lag(year),
         year_f = year,
         vha_i = lag(vha),
         vha_f = vha,
         vha_change = (vha_f - vha_i)/(year_f -year_i),
         vha_harvested_yr = vha_harvested / (year_f - year_i),
         vha_yield = vha_change +  vha_harvested_yr,
         n = row_number())%>%
  ungroup()%>%
  arrange(block,year)%>%
  filter(year_i >= 1992 & n > 1)

t2b<-t2a%>%
  group_by(block_area_ha,block)%>%
  summarise(vha_change = mean(vha_change),
            vha_harvested_yr = mean(vha_harvested_yr),
            vha_yield = mean(vha_yield))

# lines to compute weighted averages
# weighted.mean(x=t2b$vha_change,w=as.numeric(t2b$block_area_ha))
# weighted.mean(x=t2b$vha_harvested_yr,w=as.numeric(t2b$block_area_ha))
# weighted.mean(x=t2b$vha_yield,w=as.numeric(t2b$block_area_ha))

t2<-t2a%>%
  select(block, year_i, year_f, vha_i, vha_f, vha_harvested_yr, vha_yield)%>%
  mutate(tx_arvested = vha_harvested_yr/vha_yield)%>%
  mutate_at(c("vha_i","vha_f","vha_harvested_yr","vha_yield"), ~round(.,digits=1))%>%
  mutate(tx_arvested = round(tx_arvested,digits=2))

colnames(t2) <- c("Block","Year_i","Year_f","Vol_i (m³/ha)","Vol_f (m³/ha)","Vol_cut (m³/ha/an)","Yield (m³/ha/an)","Harvest rate")

datatable(t2, caption = "Evolution of standing and cut volume (volume of stems and branches up to a cross-section of 7 cm in diameter) across inventory periods and forest blocks. The initial and final years of each inventory period are given in column yeari and yearf.")

wb$add_worksheet("t2_vha")
wb$add_data("t2_vha", t2, row_names = FALSE, na = "")
```

```{r}
t2a<-dendro_block%>%
  filter(year >= 1992 )%>%
  as_tibble()%>%
  select(-geom)%>%
  select(block,block,block_area_ha,year,nha,gha,vha, vcha)%>%
  left_join(harvested,join_by(block,year==year_f))%>%
  mutate(vha_harvested = volume/block_area_ha,
         vha_harvested = ifelse(is.na(vha_harvested),0,vha_harvested),
         vm_harvested = net_2019/block_area_ha,
         vm_harvested = ifelse(is.na(vm_harvested),0,vm_harvested))%>%
  group_by(block)%>%
  arrange(year)%>%
  summarise(year_i = first(year),
         year_f = last(year),
         vha_i = first(vha), # volume
         vha_f = last(vha),
         vm_i = first(vcha), # stumpage value
         vm_f= last(vcha),
         vha_harvested = sum(vha_harvested),
         vm_harvested = sum(vm_harvested))%>%
  mutate(vha_change = (vha_f - vha_i)/(year_f -year_i),
         vha_harvested_yr = vha_harvested / (year_f - year_i),
         vha_yield = vha_change +  vha_harvested_yr,
         tx_harvested_vol = vha_harvested_yr/vha_yield,
         vm_change = (vm_f - vm_i)/(year_f -year_i),
         vm_harvested_yr = vm_harvested / (year_f - year_i),
         vm_yield = vm_change +  vm_harvested_yr,
         tx_harvested_vm = vm_harvested_yr/vm_yield)

wb$add_worksheet("t2_vha_global")
wb$add_data(sheet = "t2_vha_global", t2a, row_names = FALSE, na = "")
```

```{r}
t2a%>%
  mutate(harvestToCapital_volume = vha_harvested/vha_i,
         harvestToCapital_value =  vm_harvested/vm_i,
         harvestAndIncrementToCapital_value =  (vm_harvested+(vm_f - vm_i))/vm_i)%>%
  select(block,harvestToCapital_volume,harvestToCapital_value,harvestAndIncrementToCapital_value )%>%
  mutate(across(where(is.numeric), ~ round(.x, 2))) %>%
  datatable(caption = "Turnover periods")
```

## Financials
### Cost 

```{r}
dm<-cashflows%>%
  filter(year >= 1992)%>%
  filter(!is.na(net_2019_ha) & net_2019_ha<0)%>%
  filter(category != "Estate purchase")%>%
  filter(category != "Standing value")%>%
  left_join(select(estate_purchase,
                   block,purchaseYear = year), 
            by = join_by(block))%>%
  mutate(period = 2019 - ifelse(purchaseYear>1992,purchaseYear,1992)  + 1)%>%
  group_by(block, block_name, category, period)%>%
  summarise(
    annual_net_2019_ha = sum(net_2019_ha)/period[1] )%>%
  mutate(isIncome = "Expense")

# fpalette<- c("#999999", "#E69F00", "#009E73", "#D55E00", "#CC79A7") #if more colors needed
fpalette<- rev(c("#999999", "#009E93", "#CC99A7","#9999FF"))

g3outflows<-dm%>%
  mutate(category = factor(category, levels = 
                                c("Vehicle","Tax","Road","Staff"),
                              order=T))%>%
  ggplot(aes(y=-annual_net_2019_ha,x=block_name,fill=category))+
  geom_bar(stat = "identity")+ylab("Mean annual cost (€/ha/year)")+
  theme(axis.title.x = element_blank(),legend.position = "top") + 
  scale_fill_manual(values=fpalette, name ="")+
  scale_y_continuous(limits = c(0, 550))
```

```{r}
dm%>%
  group_by(block)%>%
  summarise(annual_net_2019_ha = sum(annual_net_2019_ha))%>%
  mutate_if(is.numeric, ~round(.,digits=1))%>%
  datatable(caption = "Total expense in each block")

dm%>%
  mutate_if(is.numeric, ~round(.,digits=1))%>%
  select(-block_name, -isIncome)%>%
  datatable(caption = "Expenses by block and category")

dm%>%
  group_by(block)%>%
  mutate(tot = sum(annual_net_2019_ha))%>%
  ungroup()%>%
  mutate(prop = annual_net_2019_ha/tot)%>%
  filter(category %in% c("Staff"))%>%
  arrange(category, prop)%>%
  select(-block_name, -isIncome)%>%
  mutate_if(is.numeric, ~round(.,digits=2))%>%
  datatable(caption = "Staff expenses by block and its proportion among total expenses")

dm%>%
  group_by(block)%>%
  mutate(tot = sum(annual_net_2019_ha))%>%
  ungroup()%>%
  mutate(prop = annual_net_2019_ha/tot)%>%
  filter(category %in% c("Road"))%>%
  arrange(category, prop)  %>%
  select(-block_name, -isIncome)%>%
  mutate_if(is.numeric, ~round(.,digits=2))%>%
  datatable(caption = "Road expenses by block and its proportion among total expenses")

dm%>%
  group_by(block)%>%
  mutate(tot = sum(annual_net_2019_ha))%>%
  ungroup()%>%
  mutate(prop = annual_net_2019_ha/tot)%>%
  filter(category %in% c("Tax"))%>%
  arrange(category, prop)%>%
    select(-block_name, -isIncome)%>%
  mutate_if(is.numeric, ~round(.,digits=2))%>%
  datatable(caption = "Tax expenses by block and its proportion among total expenses")
```

```{r}
areatot<-as.numeric(sum(block_sf$block_area_ha))

mean.expenses<-dm%>%
  left_join(as_tibble(block_sf),by = join_by(block))%>%
  group_by(category)%>%
  summarise(montant_estate = sum(annual_net_2019_ha * as.numeric(block_area_ha) / areatot) * -1)

mean.expenses.plot<-ggplot(data=mean.expenses,aes(x="",y=montant_estate, fill = category))+
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) +
  scale_fill_manual(values=fpalette)
  
wb$add_worksheet("mean_expenses")
wb$add_data(sheet = "mean_expenses", mean.expenses, row_names = FALSE, na = "")
```

### Revenue
```{r}
fpalette<- c("#CC79A7", "#56B4E9", "#E69F00",  "#D55E00",  "#009E73")

rm<- cashflows%>%
  mutate(category_2 = ifelse(category == "Timber sales",
                            ifelse(product == "Firewood",
                                   "Firewood","Log"),
                            category))%>%
  filter(!is.na(net_2019_ha) & net_2019_ha>0)%>%
  filter(category != "Standing value")%>%
  left_join(select(estate_purchase,
                   block,purchaseYear = year), 
            by = join_by(block))%>%  
  mutate(period = 2019 - ifelse(purchaseYear>1992,purchaseYear,1992)  + 1)%>%
  group_by(block, block_name, category_2)%>%
  summarise(annual_net_2019_ha = sum(net_2019_ha)/period[1])%>%
  ungroup()%>%
  mutate(isIncome = "Income")

rm%>%
    group_by(block)%>%
    summarise(annual_net_2019_ha = sum(annual_net_2019_ha))%>%
    mutate_if(is.numeric, ~round(.,digits=1))%>%
    datatable(caption = "Total revenues by block")

rm%>%
  select(-block_name, -isIncome)%>%
  mutate_if(is.numeric, ~round(.,digits=2))%>%
  datatable(caption = "Revenues by block and category")

rm%>%
  group_by(block)%>%
  mutate(tot = sum(annual_net_2019_ha))%>%
  ungroup()%>%
  mutate(prop = annual_net_2019_ha/tot)%>%
  filter(category_2 %in% c("Firewood","Log"))%>%
  arrange(category_2, prop)%>%
    select(-block_name, -isIncome)%>%
  mutate_if(is.numeric, ~round(.,digits=2))%>%
  datatable(caption = "Wood-product revenues by block and its share amont toal revenues")

rm%>%
  group_by(block)%>%
  mutate(tot = sum(annual_net_2019_ha))%>%
  ungroup()%>%
  mutate(prop = annual_net_2019_ha/tot)%>%
  filter(category_2 %in% c("Hunting"))%>%
  arrange(category_2, prop)%>%
      select(-block_name, -isIncome)%>%
    mutate_if(is.numeric, ~round(.,digits=2))%>%
  datatable(caption = "Hunting revenues by block and its share amont toal revenues")

rm%>%
  group_by(block)%>%
  mutate(tot = sum(annual_net_2019_ha))%>%
  ungroup()%>%
  mutate(prop = annual_net_2019_ha/tot)%>%
  filter(category_2 %in% c("Fishing"))%>%
  arrange(category_2, prop)%>%
      select(-block_name, -isIncome)%>%
    mutate_if(is.numeric, ~round(.,digits=2))%>%
  datatable(caption = "Fishing revenues by block and its share amont toal revenues")

rm%>%
  group_by(block)%>%
  mutate(tot = sum(annual_net_2019_ha))%>%
  ungroup()%>%
  mutate(prop = annual_net_2019_ha/tot)%>%
  filter(category_2 %in% c("Subsidy"))%>%
  arrange(category_2, prop)%>%
      select(-block_name, -isIncome)%>%
    mutate_if(is.numeric, ~round(.,digits=3))%>%
  datatable(caption = "Subsidy revenues by block and its share amont toal revenues")

```

```{r}
intial_capital<-dendro_block%>%
  as_tibble()%>%
  group_by(block)%>%
  filter(year>= 1992)%>%
  filter(year == min(year))%>%
  select(block, initial_year = year, initial_vha = vha, initial_vm = vcha)

final_capital<-dendro_block%>%
  as_tibble()%>%
  group_by(block)%>%
  filter(year == 2019)%>%
  select(block, final_year = year, final_vha = vha, final_vm = vcha)

tt<-cashflows%>%
  filter(category == "Timber sales")%>%
  filter(year >= 1992)%>%
  mutate(volume_ha = as.numeric(volume/block_area_ha))%>%
  left_join(intial_capital, by = join_by(block))%>%
  left_join(final_capital, by = join_by(block))%>%
  group_by(block,year,initial_year,initial_vha,initial_vm,final_year,final_vha,final_vm)%>%
  summarise(volume_ha = sum(volume_ha),
            income_ha = sum(net_2019_ha))%>%
  group_by(block,initial_year,initial_vha,initial_vm,final_year,final_vha,final_vm)%>%
  arrange(year)%>%
  mutate(vol_cum = cumsum(volume_ha),
         val_cum = cumsum(income_ha))

# ggplot(data=tt, aes(y=val_cum, x=year, col=block, group=block))+geom_line()+
#   geom_hline(aes(yintercept = initial_vm, col=block))
# ggplot(data=tt, aes(y=vol_cum, x=year, col=block, group=block))+geom_line()+
#   geom_hline(aes(yintercept = initial_vha, col=block))

t9_rotation<-tt%>%
  filter(year==2018)%>%
  mutate(tx_vol = vol_cum/initial_vha,
         tx_val = val_cum/initial_vm,
          tx_rotation_vol = (vol_cum + (final_vha - initial_vha))/initial_vha,
         tx_rotation_val = (val_cum + (final_vm - initial_vm))/initial_vm) 
         
wb$add_worksheet("t9_rotation")
wb$add_data(sheet = "t9_rotation", t9_rotation, row_names = FALSE, na = "")

t9_rotation%>%
  ungroup()%>%
  mutate_if(is.numeric, ~round(.,digits=3))%>%
  datatable(caption = "Estimating the ratio of the volume/value of timber produced over the initial volume/value.")
```

### Net revenue
```{r}
t9_bilan<-rm%>%
  rename(category = category_2)%>%
  bind_rows(dm)%>%
  ungroup()%>%
  select(block, category, value = annual_net_2019_ha, isIncome)

wb$add_worksheet("t9_bilan")
wb$add_data(sheet = "t9_bilan", t9_bilan, row_names = FALSE, na = "")


t10_bilanTot<-t9_bilan%>%
  group_by(block,isIncome)%>%
  summarise(val = sum(value))%>%
  pivot_wider(id_cols = block, names_from = isIncome, values_from = val)%>%
  mutate(rd = Income/Expense,
         net_revenue = Income + Expense)

wb$add_worksheet("t10_bilanTot")
wb$add_data(sheet = "t10_bilanTot", t10_bilanTot, row_names = FALSE, na = "")

t10_bilanTot%>%
      mutate_if(is.numeric, ~round(.,digits=2))%>%
      datatable(colnames = c("Block","Expense","Revenue","Revenue/Expense ratio","Net revenue"),
                caption = "Expenses and revenues for each block and category")
```


```{r}
g4income<-rm%>%
  mutate(category_2 = factor(category_2, levels = c("Subsidy","Fishing","Hunting","Firewood","Log"), order=T))%>%
  ggplot(aes(y=annual_net_2019_ha,x=block_name,fill=category_2))+
  geom_bar(stat = "identity")+ylab("Mean annual revenue (€/ha/year)")+
  theme(axis.title.x = element_blank(),legend.position = "top") +
  scale_fill_manual(values=fpalette, name ="")+
  scale_y_continuous(limits = c(0, 550))

flows<-ggarrange(g3outflows, g4income, labels = c("A", "B"), ncol = 2, nrow = 1,
                 common.legend = FALSE)

ggsave(flows,file = "fig/fig3.png", width=10, height = 5)
flows
```

## Profitability of the observed management strategy


```{r}
tmp<-dm%>%
  ungroup()%>%
  group_by(block)%>%
  summarise(tot = sum(annual_net_2019_ha))

net<-rm%>%
  bind_rows(dm)%>%
  group_by(block)%>%
  summarise(tot = sum(annual_net_2019_ha))
```

## Profitability indicators
### Net present value
```{r}
sev_as_cashflow<- filter(cashflows, category == "Standing value")%>%
  mutate(category = "SEV", # add soil/land expectation value in the cashflows
         net_2019_ha  = ifelse(net_2019<0, -soil_expectation_value, soil_expectation_value),
         net_2019 = net_2019_ha * as.numeric(block_area_ha))

t3<-cashflows%>%
  bind_rows(sev_as_cashflow)%>%
  mutate(category = ifelse(category == "SEV", "Capital opportunity cost", category))%>%
  mutate(category_2 = ifelse(category == "Timber sales",
                               ifelse(product == "Firewood",
                                      "Firewood","Log"),
                               category))%>%
  select(-geom)%>%
  filter(category != "Estate purchase")%>%
  left_join(select(estate_purchase,
                   block,purchaseYear = year), 
            by = join_by(block))%>% 
  mutate(initialYear = ifelse(purchaseYear>1992,purchaseYear,1992))%>%
  filter((category == "Vehicle" | year >= initialYear) & net_2019_ha != 0)%>%
  group_by(block,block_name)%>%
  mutate(year = ifelse(category == "Standing value" & net_2019_ha<0,initialYear,year))%>%
  mutate(year2 = year-initialYear)%>% 
  mutate(va = net_2019_ha/(1+r)^(year-initialYear))%>%
  group_by(block, block_name, category, category_2)%>%
  summarise(van_ha = sum(va))%>%
  ungroup()%>%
  select(-block)%>%
  pivot_wider(names_from = block_name, values_from = van_ha,values_fill = 0)%>%
  arrange(`Block 2`)%>%
  bind_rows(summarise(.,
                      across(where(is.numeric), sum),
                      across(where(is.character), ~"Total")))%>%
  mutate(across(where(is.numeric), ~round(.x, digits = 0)))

wb$add_worksheet("t3_npv")
wb$add_data(sheet = "t3_npv", t3, row_names = FALSE, na = "")

t3%>%
  datatable(caption = "Net present value (€/ha) per category and per forest block applying a discount rate of 2%. The first part of the table includes revenues and costs whereas the second part includes the opportunity cost of the initial capital taking into account changes in stumpage value.")
```

### Internal rate of return  
```{r}
get_irr<-function(df){
  years<-df[["year2"]]
  nets<-df[["net_2019_ha"]]
  tir<-TIRF(year = years, net = nets, fonds = soil_expectation_value, r_step = 0.0005)
  return(tir)
}

t4_all<-cashflows%>%
  select(-geom)%>%
  filter(category != "Estate purchase")%>%
    left_join(select(estate_purchase, block,purchaseYear = year),
              by = join_by(block))%>%
  mutate(initialYear = ifelse(purchaseYear>1992,purchaseYear,1992))%>%
  filter((category == "vehicle" | year >= initialYear) & net_2019_ha != 0)%>%
  group_by(block)%>%
  mutate(year = ifelse(category == "Standing value" & net_2019_ha<0,initialYear,year))%>%
  mutate(year2 = year-initialYear)%>% 
  select(block, year, year2, initialYear, net_2019_ha, category)%>%
  nest()%>%
  mutate(tir = map_dbl(data,get_irr),
         n = map_dbl(data, ~ max(.$year2))) %>%
  select(block, n, tir)%>%
  mutate(TIR_nominal = (tir+1)*(1+inflation_rate)-1)

wb$add_worksheet("t4_tir_all")
wb$add_data(sheet = "t4_tir_all", t4_all, row_names = FALSE, na = "")

t4_all%>%
    arrange(block)%>%
  mutate_if(is.numeric, ~round(.,digits=3))%>%
  datatable(caption = "Estimates of the internal rate of return expressed in real values (IRR) or in nominal values (IRRnominal) and taking into account all financial flows ")
  
```


```{r}
t4_comp<-cashflows%>%
  select(-geom)%>%
  filter(category != "Estate purchase")%>%
    filter(!category %in% c("Road","vehicle","Hunting","Fishing","Subsidy","Tax"))%>%
    left_join(select(estate_purchase,    
                     block,purchaseYear = year),   
              by = join_by(block))%>%
  mutate(initialYear = ifelse(purchaseYear>1992,purchaseYear,1992))%>%
  filter((category == "vehicle" | year >= initialYear) & net_2019_ha != 0)%>%
  group_by(block)%>%
  mutate(year = ifelse(category == "Standing value" & net_2019_ha<0,initialYear,year))%>%
  mutate(year2 = year-initialYear)%>% 
  select(block, year, year2, initialYear, net_2019_ha, category)%>%
  nest()%>%
  mutate(tir = map_dbl(data,get_irr),
         n = map_dbl(data, ~ max(.$year2))) %>%
  select(block, n, tir)%>%
  mutate(TIR_nominal = (tir+1)*(1+inflation_rate)-1)

wb$add_worksheet("t4_tir_comp")
wb$add_data(sheet = "t4_tir_comp", t4_comp, row_names = FALSE, na = "")

t4_comp%>%
  mutate_if(is.numeric, ~round(.,digits=3))%>%
  arrange(block)%>%
  datatable(caption = "Estimates of the internal rate of return expressed in real values (IRR) or in nominal values (IRRnominal) and excluding those concerning roadworks, subsidies, taxes, fishing, hunting.")

```

```{r}
# This code chunk generates a summary table that will later be merged with simulation results
# only the outflows that can be compared to simulation results are kept (staff, silvicultural interventions)
tmp_d<-dm%>%
  filter(category == "Staff")%>%
  group_by(block)%>%
  summarise(dm = sum(annual_net_2019_ha))%>%
  select(block, dm)

tmp_r<-rm%>%
  filter(category_2 %in% c("Log","Firewood"))%>%
  group_by(block)%>%
  summarise(rm = sum(annual_net_2019_ha))%>%
  select(block, rm)

tmp_npv1<-t3%>%
  filter(category_2 %in% c("Staff","Firewood","Log"))%>%
  select(-category,-category_2)%>%
  summarise_all(.funs = sum)%>%
  pivot_longer(everything(), cols = , names_to = "block", values_to = "npv1")%>%
  mutate(block = as.numeric(str_sub(block, start = -1, end = -1)))

tmp_npv4<-t3%>%
  filter(category_2 %in% c("Staff","Firewood","Log","Capital opportunity cost","Standing value"))%>%
  select(-category,-category_2)%>%
  summarise_all(.funs = sum)%>%
  pivot_longer(everything(), cols = , names_to = "block", values_to = "npv4")%>%
  mutate(block = as.numeric(str_sub(block, start = -1, end = -1)))

tmp_irr<- t4_comp%>%
  ungroup()%>%
  select(block, irr5 = tir)%>%
  mutate(nominal_irr = (irr5+1)*(1+inflation_rate)-1)


sumryTable<-t2a%>%
  mutate(nb_year = year_f - year_i,
         vt = vha_harvested,
         vm = vt / nb_year)%>%
  select(block, nb_year, vm, vt, yield = vha_yield)%>%
  left_join(tmp_d, by = join_by(block))%>%
  mutate(dt =dm*nb_year)%>%
  left_join(tmp_r, by = join_by(block))%>%
  mutate(rt = rm*nb_year)%>%
  left_join(tmp_npv1, by = join_by(block))%>%
  left_join(tmp_npv4, by = join_by(block))%>%
  left_join(tmp_irr, by = join_by(block))%>%
  mutate(npv_inf = npv1 * (1+discount_rate)^nb_year/(((1+discount_rate)^nb_year-1)))%>%
  mutate(source = "Field data",
         scenario = paste0("Block ",block),
         species = "mixed",
         hazard = "observation")%>%
  select(-block, -vt, -yield)
```

## Profitability of the simulated alternative management scenarios


```{r}
load("data/simulations/DOU.rdata")
load("data/simulations/SPR.rdata")
load("data/simulations/OAK.rdata")
load("data/simulations/MIX.rdata")

# simple functions
totalIncome<-function(net){
  mi<-sum(ifelse(net>0,net,0))
  return(mi)
}
totalExpense<-function(net){
  mi<-sum(ifelse(net<0,net,0))
  return(mi)
}
meanIncome<-function(net, rotation){
  mi<-sum(ifelse(net>0,net,0))/rotation
  return(mi)
  }
meanExpense<-function(net, rotation){
  mi<-sum(ifelse(net<0,net,0))/rotation
  return(mi)
}
volumeProduction<-function(volume, rotation){
  v<-sum(volume, na.rm = T)/rotation
  return(v)
}
dico_hazard<-tibble(label = c(paste0("DOU",1:3),
                              paste0("SPR",1:3),
                              paste0("OAK",1:3),
                              paste0("OAK_SHLTR",1:3),
                              paste0("MIX",1:3)),
                    hazard = c("Hazard-free","Growth decline","Rege. difficulties",
                               "Hazard-free","Bark beetles","Rege. difficulties",
                               "Hazard-free","Growth decline","Rege. difficulties",
                               "Hazard-free","Growth decline","Rege. difficulties",
                               "Hazard-free","Growth decline","Ash dieback"))

```

```{r}

t5<-DOU%>%
  bind_rows(SPR)%>%
  bind_rows(OAK)%>%
  bind_rows(MIX)%>%
  mutate(rm = map2_dbl(billbook,rotation, ~ meanIncome(.x$net,.y)))%>%
  mutate(dm = map2_dbl(billbook,rotation, ~ meanExpense(.x$net,.y)))%>%
  mutate(dt = map_dbl(billbook, ~ totalExpense(.x$net)))%>%
  mutate(rt = map_dbl(billbook, ~ totalIncome(.x$net)))%>%
  mutate(vm = map2_dbl(billbook,rotation, ~ volumeProduction(.x$volume, .y)))%>%
  mutate(irr5 = irr3)%>%
  mutate(nominal_irr = (irr5+1)*(1+inflation_rate)-1)%>%
  mutate(npv_inf = npv1 * (1+discount_rate)^rotation / ((1+discount_rate)^rotation - 1))%>%
  left_join(dico_hazard, by = join_by(label))%>%
  select(scenario = label, species, hazard, nb_year = rotation, 
         rt, dt, rm, dm, vm, npv1, npv1_n_fr = npv_n_fr, npv4, npv_inf, npv_inf_fr, irr5, nominal_irr)%>%
  mutate(source = "Simulation")%>%
  select(source, scenario:nominal_irr)%>%
  bind_rows(sumryTable)

wb$add_worksheet("t5_comparScenarios")
wb$add_data(sheet = "t5_comparScenarios", t5, row_names = FALSE, na = "")

t5%>%
  mutate(across(rt:npv_inf_fr, ~ round(.x, digits=0)))%>%
  mutate(across(irr5:nominal_irr, ~ round(.x, digits=3)))%>%
  datatable(caption = "Profitability indicators of the alternative simulated scenarios.")

```


```{r}
t<-DOU%>%
  bind_rows(SPR)%>%
  bind_rows(OAK)%>%
    mutate(max_stumpage = map_dbl(billbook, ~ max(.x$net)))%>%
  filter(label %in% c("DOU1","SPR1","OAK1"))%>%
  select(label, rotation, max_stumpage)
```

```{r}
intial_capital%>%
  left_join(final_capital,by="block")%>%
  ungroup()%>%
    mutate(period = final_year - initial_year)%>%
  mutate(stumpage_change_pc = (final_vm-initial_vm)/initial_vm *100)%>%
  select(block, period, stumpage_change_pc)%>%
  mutate_if(is.numeric, ~round(.,digits=1))%>%
  datatable(caption = "Evolution of the stumpage value across the blocks.", 
            colnames = c("Block","nb. years","% stumpage change"))
```


```{r}
MIX_standevolution%>%
  filter(status == "alive")%>%
  filter(year == min(year) | year == max(year))%>%
  group_by(label)%>%
  arrange(year)%>%
  summarise(initial_volume = first(totVolume),
            final_volume = last(totVolume),
            change_volume = (final_volume-initial_volume)/initial_volume*100,
            initial_value = first(totValue),
            final_value = last(totValue),
            change_value = (final_value-initial_value)/initial_value*100)%>%
    mutate_if(is.numeric, ~round(.,digits=1))%>%
    datatable(caption = "Evolution of stand volume (V) and stumpage value (S) under the MIX simulated scenario. Initial (t=0) and final (t=32) stand volume were approximately the same whereas the stumpage value increased.",
              colnames = c("Scenario","V(t=0)","V(t=32)","% volume change",
                           "S(t=0)","S(t=32)","% stumpage change"))
```


```{r}
t1<-t5%>%
  select(source, scenario, nb_year, rt, rm, dt, dm,vm,npv1,npv4, npv_inf, npv_inf_fr, irr5)%>%
  mutate(rt=rt/1000,
         rm =rm/1000,
          dt = -dt/1000,
         dm = -dm/1000,
         npv1 = npv1/1000,
         npv4 = npv4/1000,
         npv_inf = npv_inf/1000,
         npv_inf_fr = npv_inf_fr/1000,
         irr5 = irr5 * 100)%>%
  mutate(subscenario = if_else(source == "Simulation", 
                               str_sub(scenario, start = -1, end = -1),
                               NA))%>%
  mutate(label = if_else(source == "Simulation", 
                               str_sub(scenario, start = 1, end = -2),
                               "Field data"))%>%
  pivot_longer(cols = nb_year:irr5,names_to = "var")%>%
  mutate(label = factor(label, levels = c("DOU","SPR","OAK","OAK_SHLTR","MIX","Field data"), 
                        labels = c("DOU","SPR","OAK","OAK SHLTR","MIX","Field data"),
                        ordered= T))

t2_errorbar<-t1%>%
  filter(label != "Field data")%>%
  group_by(label,var)%>%
  summarise(min = min(value),
            max = max(value))

mySimulationPlot<-function(variable, variable_name, uncomparable_scenario = NULL){
  mycolors<-tibble(label = levels(t1$label),
                   col = brewer.pal(6, "Dark2"))%>%
    mutate(col2 = if_else(label %in% uncomparable_scenario, "grey80", col))
  
  g<-ggplot()+
    geom_point(data=filter(t1,(is.na(subscenario)  & var == variable) | (subscenario == "1" & var == variable)), 
               aes(x=label,y=value, col=label),
               size=2)+
    geom_errorbar(data = filter(t2_errorbar,var == variable), 
                  aes(x=label, y = max, ymin=min, ymax=max, col=label),width=0,linewidth = 1)+
    ylab(variable_name) +
    xlab(NULL) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "none",  # removed here, combined later
      axis.title.y = element_text(size = 8),
      axis.text.y = element_text(size = 7),
      axis.text.x = element_text(size = 7,angle = 90,hjust = 1),
      panel.grid.major.x = element_line(color = "grey85"),
      panel.grid.minor = element_blank(),
      panel.border = element_rect(color = "grey60", fill = NA, linewidth = 0.4),
      plot.margin = margin(2, 2, 2, 2)  
    )+
    scale_color_manual(values = mycolors$col2)                      
  
  return(g)
  
}


g1<-mySimulationPlot("nb_year", "Study period (years)")
g2<-mySimulationPlot("vm", "Harvested volume (m³/ha/year)")
g3<-mySimulationPlot("rt", "Total revenue over the study period (k€/ha)")
g4<-mySimulationPlot("rm", "Revenue (k€/ha/year)")
g5<-mySimulationPlot("dt", "Total expense over the study period (k€/ha)")
g6<-mySimulationPlot("dm", "Expense (k€/ha/year)")
# g7<-mySimulationPlot("van1", expression(NPV[n] ~ " excl. captial opportunity cost (k€/ha)"))
# g8<-mySimulationPlot("van4", expression(NPV[n] ~ " incl. captial opportunity cost (k€/ha)"))
g9<-mySimulationPlot("npv_inf", expression(NPV[infinity] ~ "(k€/ha)"), uncomparable_scenario = c("MIX","OAK SHLTR"))
g10<-mySimulationPlot("irr5", "Internal rate of return (%)", uncomparable_scenario = c("MIX","OAK SHLTR"))
g11<- mySimulationPlot("npv_inf_fr", expression("Fully regulated" ~ NPV[infinity] ~ "(k€/ha)"))
  
gall<-ggarrange(g1,g2,g4,g6,g9,g11,g10, ncol = 4, nrow = 2,
                labels = c("A","B","C","D","E","F","G"))
gall
ggsave(gall, filename = "fig/fig4.png", width = 8, height = 7)

```


# Supplementary material

```{r}
fwd<-data.frame(circ = unique(price_list$circ), species_group_en = "Firewood", price_2019 = firewood_price)
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7","black")
ggprice<-price_list%>%
  bind_rows(fwd)%>%
  mutate(species_group_en = factor(species_group_en , levels=c(sort(unique(price_list$species_group_en)),"Firewood"), order=T))%>%
  ggplot(aes(y=price_2019,x=circ/pi, col =  species_group_en))+geom_line()+
  xlab("Diameter (cm)")+ylab("Price (€/m³)")+
  theme(legend.title = element_blank())+
  scale_color_manual(values=cbPalette)
ggprice

ggsave(file="fig/suppl_fig1.png", ggprice, width=6, height=4)
```

```{r}
t7_achat<-estate_purchase%>%
  mutate(price = - net_2019)%>%
  left_join(block_sf,by = join_by(block,block_name))%>%
  mutate(block_area_ha = as.numeric(block_area_ha))%>%
  select(block, block_name, category, price, block_area_ha)

wb$add_worksheet("t7_purchase")
wb$add_data(sheet = "t7_purchase", t7_achat, row_names = FALSE, na = "")

t8<-dendro_block%>%
  as_tibble()%>%
  select(block, year, nha, gha, vha, vcha, ghapc_oak, ghapc_conif)%>%
  mutate_at(c("ghapc_oak","ghapc_conif"), function(x) {x *100})%>%
  mutate_if(is.numeric, ~round(.,digits=1))

names(t8)<-c("Massif", "Année","NHA (/ha)","GHA (m²/ha)", "VHA (m³/ha)", "VC (€/ha)", "%Oak", "%Coniferous")

wb$add_worksheet("t8_inv")
wb$add_data(sheet = "t8_inv", t8, row_names = FALSE, na = "")
```

```{r}
evo1<-dendro_zone%>%
  ggplot(aes(x=year,y=gha))+geom_line(col="grey", aes(group=zone_name))+
  geom_line(data=dendro_block)+
  facet_grid(~block_name)+xlab("Year")+ylab("Basal area (m²/ha)")+
  theme(legend.position = "none")+
  theme(axis.text.x=element_text(angle = 45, hjust = 1,vjust = 1))+
  geom_vline(data=keydates, aes(xintercept = purchaseYear), col="orange")+
  geom_vline(data=keydates, aes(xintercept = firstInventory), col="darkorange",lty=2)
ggsave(evo1,file = "fig/supplFig2.png", width=10, height = 4)
evo1

evo2<-dendro_zone%>%
  ggplot(aes(x=year,y=vha))+geom_line(col="grey", aes(group=zone_name))+
  geom_line(data=dendro_block)+
  facet_grid(~block_name)+xlab("Year")+ylab("Volume (m³/ha)")+
  theme(legend.position = "none")+
  theme(axis.text.x=element_text(angle = 45, hjust = 1,vjust = 1))+
  geom_vline(data=keydates, aes(xintercept = purchaseYear), col="orange")+
  geom_vline(data=keydates, aes(xintercept = firstInventory), col="darkorange",lty=2)

ggsave(evo2,file = "fig/supplFig3.png", width=10, height = 4)
evo2

evo3<-dendro_zone%>%
  ggplot(aes(x=year,y=vcha))+geom_line(col="grey", aes(group=zone_name))+
  geom_line(data=dendro_block)+
  facet_grid(~block_name)+xlab("Year")+ylab("Value (€/ha)")+
  theme(legend.position = "none")+
  theme(axis.text.x=element_text(angle = 45, hjust = 1,vjust = 1))+
  geom_vline(data=keydates, aes(xintercept = purchaseYear), col="orange")+
  geom_vline(data=keydates, aes(xintercept = firstInventory), col="darkorange",lty=2)

ggsave(evo3,file = "fig/supplFig4.png", width=10, height = 4)
evo3

evo5<-dendro_zone%>%
  ggplot(aes(x=year,y=ghapc_oak))+geom_line(col="grey", aes(group=zone_name))+
  geom_line(data=dendro_block)+
  facet_grid(~block_name)+xlab("Year")+ylab("%oak basal area")+
  theme(legend.position = "none")+
  theme(axis.text.x=element_text(angle = 45, hjust = 1,vjust = 1))+
  geom_vline(data=keydates, aes(xintercept = purchaseYear), col="orange")+
  geom_vline(data=keydates, aes(xintercept = firstInventory), col="darkorange",lty=2)

ggsave(evo5,file = "fig/supplFig5.png", width=10, height = 4)
evo5

evo4<-dendro_zone%>%
  ggplot(aes(x=year,y=n_sp))+geom_line(col="grey", aes(group=zone_name))+
  geom_line(data=dendro_block)+
  facet_grid(~block_name)+xlab("Year")+ylab("Nb of species")+
  theme(legend.position = "none")+
  theme(axis.text.x=element_text(angle = 45, hjust = 1,vjust = 1))+
  geom_vline(data=keydates, aes(xintercept = purchaseYear), col="orange")+
  geom_vline(data=keydates, aes(xintercept = firstInventory), col="darkorange",lty=2)

ggsave(evo4,file = "fig/supplFig6.png", width=10, height = 4)
evo4

evo6<-dendro_zone%>%
  ggplot(aes(x=year,y=vcha/vha))+geom_line(col="grey", aes(group=zone_name))+
  geom_line(data=dendro_block)+
  facet_grid(~block_name)+xlab("Year")+ylab("Value / Volume (€/m³)")+
  theme(legend.position = "none")+
  theme(axis.text.x=element_text(angle = 45, hjust = 1,vjust = 1))+
  geom_vline(data=keydates, aes(xintercept = purchaseYear), col="orange")+
  geom_vline(data=keydates, aes(xintercept = firstInventory), col="darkorange",lty=2)

ggsave(evo6,file = "fig/supplFig7.png", width=10, height = 4)
evo6
```

```{r}
# export billbooks of all scenarios (excluding management/staff cost)
scenarios<-DOU%>%
  bind_rows(SPR)%>%
  bind_rows(OAK)%>%
  bind_rows(MIX)


for(i in 1:nrow(scenarios)){
  name<-paste(scenarios[i,"species"],"--", scenarios[i,"label"])
  print(name)
  
  bb<-scenarios[i,]%>%
    select(billbook)%>%
    unnest(cols = c(billbook))%>%
    select(-isIncome)%>%
    mutate_at(c("value","volume","net"), ~round(.,digits=0))%>%
    filter(category != "Management")%>%
    select(label, year, value, volume, net)
  
  names(bb)<-c("Label", "Year","Price (€/ha)","Volume (m³)", "Net (€/ha)")
  
  bb%>%
    datatable(caption = paste0("Billbook of ", scenarios[i,"label"], " excluding annual staff expenses."))%>%
    print()
  
  wb$add_worksheet(name)
  wb$add_data(sheet = name, bb, row_names = FALSE, na = "")
}
```

```{r}
g1<-MIX_standevolution%>%
  filter(status == "alive")%>%
  ggplot(aes(x=date,y=totVolume,col=label))+geom_line()+
  ylab("Volume (m³/ha)")+xlab("Year")+
  labs(col = "Scenario")+
  theme(legend.position="top")

g2<-MIX_standevolution%>%
  filter(status == "alive")%>%
  mutate(totValue = totValue/1000)%>%
  ggplot(aes(x=date,y=totValue,col=label))+geom_line()+
  ylab("Stumpage value (k€/ha)")+xlab("Year")+
  labs(col = "Scenario")+
  theme(legend.position="top")

g12<-ggarrange(g1,g2, ncol = 2, nrow = 1,
                labels = c("A","B"))
g12
ggsave(g12, filename = "fig/supplFig8.png", width = 7, height = 3)
```




```{r}
#save workbook
wb$save("tab/tables.xlsx")
```

