---
title: "Sensitivity analyses on land expectation value."
author: "Gauthier Ligot"
date: "2026-07-17"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(sf)
library(mapview)
library(DT)
library(knitr)
library(ggpubr)
library(RColorBrewer)
library(openxlsx2)
library(kableExtra)
```

# loading data as in paper_result_v7
```{r}
source("../script/tirf.R") # a function to compute the internal rate of return
source("../script/myGGplotConfigs.R") # settings for personnalized ggplot layouts
load("../data/1_CalcData_en.Rdata")  # inventory data
load("../data/3_calcRevenues_en.Rdata") # economic data
source("../script/commonParameters.R") # economic parameter
fpalette<- c("#E69F00", "#999999", "#009E73", "#D55E00", "#CC79A7","#333399") # chosen colors for the plots
```

# Sensitivity analysis one one forest bloc (block 2)
I computed NPV2 with $L$ ranging from 1000 to 15000. 


```{r echo = TRUE}
# L <- 5000
L_step <- 1000
L_min <- 1000
L_max <- 15000
selected_block <- 2 # I selected abitrarily the largest block
```

```{r include = FALSE}
options(dplyr.summarise.inform = FALSE)

# Build the appropriate billbook
billbook<-cashflows%>%
  filter(block == selected_block)%>% # the selected block
  mutate(category_2 = ifelse(category == "Timber sales",
                             ifelse(product == "Firewood",
                                    "Firewood","Log"),
                             category))%>%
  filter(category_2 %in% c("Staff","Firewood","Log","Standing value"))%>% # keep the same cash flows as in table 5
  select(-geom)%>%
  filter(category != "Estate purchase")%>%
  left_join(select(estate_purchase,
                   block,purchaseYear = year), 
            by = join_by(block))%>% 
  mutate(initialYear = ifelse(purchaseYear>1992,purchaseYear,1992))%>% 
  filter(((category == "Vehicle" | year >= initialYear) & net_2019_ha != 0))%>% # cashflows of 0€ or outside the study period
  # group_by(block,block_name)%>% # not needed, here we work for only one block
  mutate(year = ifelse(category == "Standing value" & net_2019_ha<0,initialYear,year))%>%
  mutate(year2 = year-initialYear)%>% 
  arrange(year2)%>%
  mutate(pv = net_2019_ha/(1+r)^(year2))%>% # present value
  select(block, category = category_2, year = year2, net_2019_ha, pv_2019_ha = pv)
```

```{r echo=TRUE}
block_output<-tibble()
for(L in seq(L_min,L_max,L_step)){
  
  n<-max(billbook$year)
  npv2 <- sum(billbook$pv_2019_ha) + L/(1+r)^n - L
  irr<-TIRF(year = billbook$year, net = billbook$net_2019_ha, fonds = L, r_step = 0.0001) #This takes time !
  
  block_output<-block_output%>%
    bind_rows(tibble(label = "Block 2", L=L,npv2=npv2,irr=irr))
}
```


It is quite obvious that $L$ affects directly and linearly NPV2. We knew it looking at its formula...

```{r echo=FALSE}
ggplot(block_output,aes(x=L,y=npv2))+geom_line()+geom_point()+xlab("Land expectation value")+ylab("NPV2")
```

It is quite obvious that $L$ affects directly and non-linearly IRR. Not new neither...
```{r echo = FALSE}
ggplot(block_output,aes(x=L,y=irr))+geom_line()+geom_point()+xlab("Land expectation value")+ylab("IRR")
```

# Sensitivity analysis for the theoretical scenarion of spruce plantation (SPR1)
```{r echo=TRUE}
load("../data/simulations/SPR.rdata")

spr_billbook<-SPR%>%
  filter(label == "SPR1")%>%
  pull(billbook)%>%
  .[[1]]%>%
  select(category, year, net_2019_ha = net)%>%
  mutate(pv_2019_ha = net_2019_ha/(1+r)^year)

spr_tmp<-tibble()

spr_output<-tibble()
for(L in seq(L_min,L_max,L_step)){
  
  n<-max(spr_billbook$year)
  npv2 <- sum(spr_billbook$pv_2019_ha) + L/(1+r)^n - L
  irr<-TIRF(year = spr_billbook$year, net = spr_billbook$net_2019_ha, fonds = L, r_step = 0.0001) #This takes time !
  
  spr_output<-spr_output%>%
    bind_rows(tibble(label="SPR1", L=L,npv2=npv2,irr=irr))
}
```

```{r echo=FALSE}
dat<-block_output%>%
  bind_rows(spr_output)

g1<-ggplot(dat,aes(x=L,y=npv2,col=label))+geom_point()+geom_line()+xlab("Land expectation value")+ylab("NPV2")
g2<-ggplot(dat,aes(x=L,y=irr,col=label))+geom_point()+geom_line()+xlab("Land expectation value")+ylab("IRR")

ggarrange(g1, g2, labels = c("A", "B"), ncol = 2, nrow = 1,common.legend = TRUE)
```


